---
title: "CAMCAN_cog"
author: "Josefina Weinerova"
date: "2025-09-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#packages
```{r}
if (!require(pacman)) install.packages("pacman")

pacman::p_load(lmerTest, tidyverse, effectsize, Rmisc, cowplot, broom.mixed,purrr, dplyr,openxlsx)
```
#load data (data available from the CamCAN team: https://cam-can.mrc-cbu.cam.ac.uk/dataset/)

```{r}
df <- read_csv("CamCAN_demographic_cognitive_COVID_P1_P4_P5.csv")
covid_q <- read_csv("just_covid_answers_2.csv")
```
#Exclusion criteria
```{r}
df$CC.ID <- as.factor(df$CC.ID)
df <- df %>%
  group_by(CC.ID) %>%
  fill(covid_question, .direction = "downup") %>% # fill down, then up within each participant
  ungroup()
df <- df %>% filter(covid_question == "No" | covid_question == "Yes, confirmed by a positive test")
df <- df %>% mutate(covid_group = as.integer(covid_question == "Yes, confirmed by a positive test"))
```
#df pre-processing
```{r}
df$covid_group <- as.factor(df$covid_group)
df <- df %>%
  mutate(wave = str_remove_all(wave, "\\s+"))
# removal of outliers

df_wide <- df %>%
  pivot_wider(
    id_cols = CC.ID,
    names_from = wave, 
    values_from = c(
      story_memory_immediate_score,
      story_memory_delayed_score,
      cattell_total_score,
      acer_score,
      stw_n_correct_trials 
    ),
    names_glue = "{.value}_{wave}" # new names: <original>_<wave>
  )

colnames(df_wide)
df_wide <- df_wide %>% select(-c("story_memory_immediate_score_Phase4", "story_memory_delayed_score_Phase4"))

df_wide$story_mem_immed_dif <- df_wide$story_memory_immediate_score_Phase5 - df_wide$`story_memory_immediate_score_Phase1/2`
df_wide$story_mem_del_dif <- df_wide$story_memory_delayed_score_Phase5 - df_wide$`story_memory_delayed_score_Phase1/2`
```

# Removing outliers
- calculates difference scores and then removes values that have difference scores that are 3x IQR above or below third and first quartile

```{r}
# removing outliers
outliers <- function(variable) {
  iqr <- IQR(variable, na.rm = TRUE)
  Q <- quantile(variable, probs = c(.25, .75), na.rm = TRUE)
  mild_low <- Q[1] - (3 * iqr)
  mild_high <- Q[2] + (3 * iqr)
  new_variable <- ifelse(variable < mild_low | variable > mild_high, NA, variable)
  return(new_variable)
}

# NA numbers
df_wide %>%
  dplyr::summarise(across(everything(), ~ sum(is.na(.))))
# creating difs columns
# Long: gather all columns ending with _Phase1/2, _Phase4, _Phase5
long <- df_wide %>%
  pivot_longer(
    cols = matches("_Phase(1/2|4|5)$"),
    names_to = c("var", "phase"),
    names_pattern = "^(.*)_Phase(1/2|4|5)$",
    values_to = "value"
  ) %>%
  mutate(phase = recode(phase, "1/2" = "12")) # avoid slash in names

# Compute diffs per participant & variable, then wide df
diffs <- long %>%
  select(CC.ID, var, phase, value) %>%
  pivot_wider(names_from = phase, values_from = value) %>% 
  mutate(
    dif1 = `5` - `4`, # Phase5 − Phase4
    dif2 = `5` - `12` # Phase5 − Phase1/2
  ) %>%
  select(CC.ID, var, dif1, dif2) %>%
  pivot_wider(
    id_cols = CC.ID,
    names_from = var,
    values_from = c(dif1, dif2),
    names_glue = "{var}_{.value}" 
  )

# attach back to original data
df2 <- df_wide %>% left_join(diffs, by = "CC.ID")

cols_difs <- c("story_memory_immediate_score_dif1", "story_memory_delayed_score_dif1", "cattell_total_score_dif1", "acer_score_dif1", "stw_n_correct_trials_dif1", "story_memory_immediate_score_dif2", "story_memory_delayed_score_dif2", "cattell_total_score_dif2", "acer_score_dif2", "stw_n_correct_trials_dif2")

na_pre <- colSums(is.na(df2))
# removing outliers
df2[cols_difs] <- sapply(df2[cols_difs], outliers) # applying outlier removal

na_post <- colSums(is.na(df2))
#check numbers pre and post removal
na_pre
na_post
```
# original columns NA if the difference is NA
```{r}
# List of test names for those that have 3 instances
test_names <- cols <- c(
  "cattell_total_score",
  "stw_n_correct_trials"
)

# Loop over each test and replace T1 and T2 with NA where the difference is NA
for (test in test_names) {
  diff_col1 <- paste0(test, "_dif1")
  diff_col2 <- paste0(test, "_dif2") # Difference column
  t1_col <- paste0(test, "_Phase1/2") # Time 1 column
  t2_col <- paste0(test, "_Phase4")
  t3_col <- paste0(test, "_Phase5") # Time 2 column

  # Set T1 and T2 to NA where the difference column is NA
  df2[[t1_col]][is.na(df2[[diff_col1]])] <- NA
  df2[[t2_col]][is.na(df2[[diff_col1]])] <- NA
  df2[[t3_col]][is.na(df2[[diff_col1]])] <- NA
  df2[[t1_col]][is.na(df2[[diff_col2]])] <- NA
  df2[[t2_col]][is.na(df2[[diff_col2]])] <- NA
  df2[[t3_col]][is.na(df2[[diff_col2]])] <- NA
}
#tests with 2 instances
test_names <- cols <- c(
  "story_memory_delayed_score",
  "story_memory_immediate_score",
  "acer_score"
)

# Loop over each test and replace T1 and T2 with NA where the difference is NA
for (test in test_names) {
  diff_col2 <- paste0(test, "_dif2") # Difference column
  t1_col <- paste0(test, "_Phase1/2") # Time 1 column
  t3_col <- paste0(test, "_Phase5") # Time 3 column

  # Set T1 and T2 to NA where the difference column is NA
  df2[[t1_col]][is.na(df2[[diff_col2]])] <- NA
  df2[[t3_col]][is.na(df2[[diff_col2]])] <- NA
}

View(df2)
colSums(!is.na(df2))
```
# demographics saving in another df

```{r}
colnames(df)
demo <- df %>% dplyr::select("CC.ID", "wave", "sex", "dob", "sex", "ethnic_group", "total_household_incomeP1", "total_household_incomeP5", "num_years_edu", "covid_question", "stw_age", "stw_date", "acer_age", "acer_date", "story_memory_age", "story_memory_date", "cattell_age", "cattell_date", "covid_group")


## LONG df
df_long <- df2 %>%
  select(CC.ID, matches("_Phase")) %>%
  pivot_longer(
    cols = -CC.ID,
    names_to = c("measure", "wave"), # wave keeps "Phase1/2", "Phase4", "Phase5", ...
    names_pattern = "^(.*)_(Phase.+)$",
    values_to = "value"
  ) %>%
  select(CC.ID, wave, measure, value)

## WIDE df (each measure becomes its own column)
df_long <- df_long %>%
  pivot_wider(
    names_from  = measure,
    values_from = value
  )
#merge with demographics
df_A <- merge(df_long, demo, by = c("CC.ID", "wave"))
str(df_A)
df_A <- df_A %>% mutate(
  across(where(is.character), ~ na_if(., "<NA>"))
)
#ethnicity binary variable
df_A$eth2 <- ifelse(df_A$ethnic_group == "White", "white", "other")
```

# memory analysis

```{r}
# subselect relevant columns
mem_im_data <- df_A %>% select(CC.ID, wave, story_memory_immediate_score, story_memory_delayed_score, covid_group, sex, ethnic_group, num_years_edu, story_memory_age, story_memory_date, eth2)
dates_wide <- mem_im_data %>%
  select(CC.ID, wave, story_memory_date) %>%
  distinct() %>% # in case of duplicates
  pivot_wider(
    names_from = wave, # e.g. "Phase1/2", "Phase4", ...
    values_from = story_memory_date,
    names_prefix = "story_memory_date_" # columns like story_memory_date_Phase1/2
  )
# calculate months difs between testing
mem_im_data <- mem_im_data %>%
  select(-story_memory_date) %>%
  left_join(dates_wide, by = "CC.ID")
mem_im_data <- mem_im_data %>%
  mutate(
    d_start = as.Date(`story_memory_date_Phase1/2`),
    d_end = as.Date(story_memory_date_Phase5),
    month_dif = time_length(interval(d_start, d_end), "months") # fractional
  )

mem_im_data %>%
  filter(wave == "Phase5") %>%
  filter(covid_group == 0) %>%
  summary()

mem_im_data$month_dif_sq <- mem_im_data$month_dif^2
# calculate age at T2
mem_im_data <- mem_im_data %>%
  dplyr::group_by(CC.ID) %>%
  dplyr::mutate(age2 = max(story_memory_age, na.rm = TRUE)) %>%
  ungroup()
# check data types
str(mem_im_data)
mem_im_data$sex <- as.factor(mem_im_data$sex)
mem_im_data$wave <- as.factor(mem_im_data$wave)
mem_im_data <- mem_im_data %>% filter(wave != "Phase4")
# scale continuois columns
cols <- c("age2", "month_dif", "month_dif_sq", "num_years_edu")
mem_im_data <- mem_im_data %>%
  dplyr::mutate(across(all_of(cols), ~ as.numeric(scale(.))))
# models
model_mem1 <- lmerTest::lmer(story_memory_immediate_score ~ covid_group * wave + age2 + sex + month_dif + month_dif_sq + num_years_edu + eth2 + (1 | CC.ID), data = mem_im_data)
summary(model_mem1)
mem1_stand <- as.data.frame(standardize_parameters(model_mem1))

model_mem2 <- lmerTest::lmer(story_memory_delayed_score ~ covid_group * wave + age2 + sex + month_dif + month_dif_sq + num_years_edu + eth2 + (1 | CC.ID), data = mem_im_data)
summary(model_mem2)
mem2_stand <- as.data.frame(standardize_parameters(model_mem2))


# residuals check
plot(resid(model_mem1))
plot(resid(model_mem2))
qqnorm(resid(model_mem1))
qqnorm(resid(model_mem2))
```
#plots

```{r}
dfwc <- summarySEwithin(mem_im_data,
  measurevar = "story_memory_immediate_score", withinvars = "wave", betweenvars = "covid_group",
  idvar = "CC.ID", na.rm = TRUE, conf.interval = .95
)


ggplot(dfwc, aes(x = wave, y = story_memory_immediate_score, group = covid_group, color = covid_group)) +
  geom_point(size = 4) +
  geom_line(size = 2) +
  geom_errorbar(aes(ymin = story_memory_immediate_score - se, ymax = story_memory_immediate_score + se), width = .1, size = 2) +
  theme_bw() +
  labs(title = "Mean scores for instance with error bars", y = "Immediate memory score", x = "instance") +
  theme_cowplot(12) +
  theme(
    plot.title = element_blank(),
    axis.title = element_text(size = 20, face = "bold"), # originally 18 → now 20
    axis.text.x = element_text(size = 19), # originally 17 → now 19
    axis.text.y = element_text(size = 19),
    legend.title = element_text(size = 18, face = "bold"), # scale up from 16
    legend.text = element_text(size = 17), # scale up from 15
    strip.text = element_text(size = 18) # if using facet_wrap
  ) +
  scale_x_discrete(labels = c("T1", "T3")) +
  scale_color_manual(values = c("#999999", "#E69F00"), labels = c("Control group", "Covid group"), name = "Group")


dfwc <- summarySEwithin(mem_im_data,
  measurevar = "story_memory_delayed_score", withinvars = "wave", betweenvars = "covid_group",
  idvar = "CC.ID", na.rm = TRUE, conf.interval = .95
)


ggplot(dfwc, aes(x = wave, y = story_memory_delayed_score, group = covid_group, color = covid_group)) +
  geom_point(size = 4) +
  geom_line(size = 2) +
  geom_errorbar(aes(ymin = story_memory_delayed_score - se, ymax = story_memory_delayed_score + se), width = .1, size = 2) +
  theme_bw() +
  labs(title = "Mean scores for instance with error bars", y = "Delayed memory score", x = "instance") +
  theme_cowplot(12) +
  theme(
    plot.title = element_blank(),
    axis.title = element_text(size = 20, face = "bold"), # originally 18 → now 20
    axis.text.x = element_text(size = 19), # originally 17 → now 19
    axis.text.y = element_text(size = 19),
    legend.title = element_text(size = 18, face = "bold"), # scale up from 16
    legend.text = element_text(size = 17), # scale up from 15
    strip.text = element_text(size = 18) # if using facet_wrap
  ) +
  scale_x_discrete(labels = c("T1", "T3")) +
  scale_color_manual(values = c("#999999", "#E69F00"), labels = c("Control group", "Covid group"), name = "Group")
```
#plots
```{r}
ggplot(mem_im_data, aes(x = wave, y = story_memory_delayed_score, color = covid_group)) +
  geom_boxplot(size = 1) +
  labs(title = "Mean scores for instance with error bars", y = "Delayed memory score", x = "instance") +
  theme_cowplot(12) +
  theme(
    plot.title = element_blank(),
    axis.title = element_text(size = 20, face = "bold"), # originally 18 → now 20
    axis.text.x = element_text(size = 19), # originally 17 → now 19
    axis.text.y = element_text(size = 19),
    legend.title = element_text(size = 18, face = "bold"), # scale up from 16
    legend.text = element_text(size = 17), # scale up from 15
    strip.text = element_text(size = 18) # if using facet_wrap
  ) +
  scale_x_discrete(labels = c("T1", "T3")) +
  scale_color_manual(values = c("#999999", "#E69F00"), labels = c("Control group", "Covid group"), name = "Group")

ggplot(mem_im_data, aes(x = wave, y = story_memory_immediate_score, color = covid_group)) +
  geom_boxplot(size = 1) +
  labs(title = "Mean scores for instance with error bars", y = "Immediate memory score", x = "instance") +
  theme_cowplot(12) +
  theme(
    plot.title = element_blank(),
    axis.title = element_text(size = 20, face = "bold"), # originally 18 → now 20
    axis.text.x = element_text(size = 19), # originally 17 → now 19
    axis.text.y = element_text(size = 19),
    legend.title = element_text(size = 18, face = "bold"), # scale up from 16
    legend.text = element_text(size = 17), # scale up from 15
    strip.text = element_text(size = 18) # if using facet_wrap
  ) +
  scale_x_discrete(labels = c("T1", "T3")) +
  scale_color_manual(values = c("#999999", "#E69F00"), labels = c("Control group", "Covid group"), name = "Group")
```


#ACER analysis
```{r}
df_A
# subselect relevant columns
acer_data <- df_A %>% select(CC.ID, wave, acer_score, covid_group, sex, ethnic_group, num_years_edu, acer_age, acer_date, eth2)
dates_wide <- acer_data %>%
  select(CC.ID, wave, acer_date) %>%
  distinct() %>% # in case of duplicates
  pivot_wider(
    names_from = wave, # e.g. "Phase1/2", "Phase4", ...
    values_from = acer_date,
    names_prefix = "acer_date_" # columns like story_memory_date_Phase1/2
  )
# calculate months difs between testing
acer_data <- acer_data %>%
  select(-acer_date) %>%
  left_join(dates_wide, by = "CC.ID")
acer_data <- acer_data %>%
  mutate(
    d_start = as.Date(`acer_date_Phase1/2`),
    d_end = as.Date(acer_date_Phase5),
    month_dif = time_length(interval(d_start, d_end), "months") # fractional
  )
acer_data %>%
  filter(wave == "Phase5") %>%
  filter(covid_group == 1) %>%
  summary()
acer_data$month_dif_sq <- acer_data$month_dif^2
# calculate age at T2
acer_data <- acer_data %>%
  dplyr::group_by(CC.ID) %>%
  dplyr::mutate(age2 = max(acer_age, na.rm = TRUE)) %>%
  ungroup()
# check data types
str(acer_data)
acer_data$sex <- as.factor(acer_data$sex)
acer_data$wave <- as.factor(acer_data$wave)
acer_data <- acer_data %>% filter(wave != "Phase4")
# scale continuous columns
cols <- c("age2", "month_dif", "month_dif_sq", "num_years_edu")
acer_data <- acer_data %>%
  dplyr::mutate(across(all_of(cols), ~ as.numeric(scale(.))))
# models
model_acer <- lmerTest::lmer(acer_score ~ covid_group * wave + age2 + sex + month_dif + month_dif_sq + num_years_edu + eth2 + (1 | CC.ID), data = acer_data)
summary(model_acer)
acer_stand <- standardize_parameters(model_acer)

# checking residuals
plot(resid(model_acer))
plot(qqnorm(resid(model_acer)))
```
# plots
```{r}
dfwc <- summarySEwithin(acer_data,
  measurevar = "acer_score", withinvars = "wave", betweenvars = "covid_group",
  idvar = "CC.ID", na.rm = TRUE, conf.interval = .95
)


ggplot(dfwc, aes(x = wave, y = acer_score, group = covid_group, color = covid_group)) +
  geom_point(size = 4) +
  geom_line(size = 2) +
  geom_errorbar(aes(ymin = acer_score - se, ymax = acer_score + se), width = .1, size = 2) +
  theme_bw() +
  labs(title = "Mean scores for instance with error bars", y = "ACER score", x = "instance") +
  theme_cowplot(12) +
  theme(
    plot.title = element_blank(),
    axis.title = element_text(size = 20, face = "bold"), # originally 18 → now 20
    axis.text.x = element_text(size = 19), # originally 17 → now 19
    axis.text.y = element_text(size = 19),
    legend.title = element_text(size = 18, face = "bold"), # scale up from 16
    legend.text = element_text(size = 17), # scale up from 15
    strip.text = element_text(size = 18) # if using facet_wrap
  ) +
  scale_x_discrete(labels = c("T1", "T3")) +
  scale_color_manual(values = c("#999999", "#E69F00"), labels = c("Control group", "Covid group"), name = "Group")
```

#boxplots
```{r}
ggplot(acer_data, aes(x = wave, y = acer_score, color = covid_group)) +
  geom_boxplot(size = 1) +
  labs(title = "Mean scores for instance with error bars", y = "ACE-R score", x = "instance") +
  theme_cowplot(12) +
  theme(
    plot.title = element_blank(),
    axis.title = element_text(size = 20, face = "bold"), # originally 18 → now 20
    axis.text.x = element_text(size = 19), # originally 17 → now 19
    axis.text.y = element_text(size = 19),
    legend.title = element_text(size = 18, face = "bold"), # scale up from 16
    legend.text = element_text(size = 17), # scale up from 15
    strip.text = element_text(size = 18) # if using facet_wrap
  ) +
  scale_x_discrete(labels = c("T1", "T3")) +
  scale_color_manual(values = c("#999999", "#E69F00"), labels = c("Control group", "Covid group"), name = "Group")
```

# Cattel analysis
```{r}
# subselect relevant columns
cattell_data <- df_A %>% select(CC.ID, wave, cattell_total_score, covid_group, sex, ethnic_group, num_years_edu, cattell_age, cattell_date, eth2)

dates_wide <- cattell_data %>%
  select(CC.ID, wave, cattell_date) %>%
  distinct() %>% # in case of duplicates
  pivot_wider(
    names_from = wave, # e.g. "Phase1/2", "Phase4", ...
    values_from = cattell_date,
    names_prefix = "cattell_date_" # columns like story_memory_date_Phase1/2
  )
# calculate months difs between testing
cattell_data <- cattell_data %>%
  select(-cattell_date) %>%
  left_join(dates_wide, by = "CC.ID")
cattell_data <- cattell_data %>%
  mutate(
    d_start = as.Date(`cattell_date_Phase1/2`),
    d_middle = as.Date(cattell_date_Phase4),
    d_end = as.Date(cattell_date_Phase5),
    month_dif1 = time_length(interval(d_start, d_end), "months"),
    month_dif2 = time_length(interval(d_middle, d_end), "months") # fractional
  )

cattell_data %>%
  filter(wave == "Phase5") %>%
  filter(covid_group == 0) %>%
  summary()


summary(cattell_data$month_dif1, na.rm = TRUE)
summary(cattell_data$month_dif2, na.rm = TRUE)

cattell_data$month_dif1_sq <- cattell_data$month_dif1^2
cattell_data$month_dif2_sq <- cattell_data$month_dif2^2
# calculate age at T2
cattell_data$age2 <- ave(cattell_data$cattell_age, cattell_data$CC.ID,
  FUN = function(x) if (all(is.na(x))) NA_real_ else max(x, na.rm = TRUE)
)
# check data types
str(cattell_data)
cattell_data$sex <- as.factor(cattell_data$sex)
cattell_data$wave <- as.factor(cattell_data$wave)

# scale continuois columns
cols <- c("age2", "month_dif1", "month_dif1_sq", "month_dif2", "month_dif2_sq", "num_years_edu")
cattell_data <- cattell_data %>% dplyr::mutate(across(all_of(cols), ~ as.numeric(scale(.))))
# cattell_data$age2 <- scale(cattell_data$age2, center = TRUE, scale=TRUE)

contrasts(cattell_data$wave)

dfwc <- summarySEwithin(cattell_data,
  measurevar = "cattell_total_score", withinvars = "wave", betweenvars = "covid_group",
  idvar = "CC.ID", na.rm = TRUE, conf.interval = .95
)


ggplot(dfwc, aes(x = wave, y = cattell_total_score, group = covid_group, color = covid_group)) +
  geom_point(size = 4) +
  geom_line(size = 2) +
  geom_errorbar(aes(ymin = cattell_total_score - se, ymax = cattell_total_score + se), width = .1, size = 2) +
  theme_bw() +
  labs(title = "Mean scores for instance with error bars", y = "Cattell score", x = "instance") +
  theme_cowplot(12) +
  theme(
    plot.title = element_blank(),
    axis.title = element_text(size = 20, face = "bold"), # originally 18 → now 20
    axis.text.x = element_text(size = 19), # originally 17 → now 19
    axis.text.y = element_text(size = 19),
    legend.title = element_text(size = 18, face = "bold"), # scale up from 16
    legend.text = element_text(size = 17), # scale up from 15
    strip.text = element_text(size = 18) # if using facet_wrap
  ) +
  scale_color_manual(values = c("#999999", "#E69F00"), labels = c("Control group", "Covid group"), name = "Group") +
  scale_x_discrete(labels = c("T1", "T2", "T3"))

cattell_data$wave <- factor(cattell_data$wave,
  levels = c("Phase1/2", "Phase4", "Phase5")
)
# dividing into 2 models

cattell_data1 <- cattell_data %>% filter(wave != "Phase4")
summary(cattell_data1)
# model1
model_cattell1 <- lmerTest::lmer(cattell_total_score ~ covid_group * wave + age2 + sex + month_dif1 + month_dif1_sq + num_years_edu + eth2 + (1 | CC.ID), data = cattell_data1)
summary(model_cattell1)
standardize_parameters(model_cattell1)
cattell1_stand <- as.data.frame(standardize_parameters(model_cattell1))
cattell_data2 <- cattell_data %>% filter(wave != "Phase1/2")
# model2
model_cattell2 <- lmerTest::lmer(cattell_total_score ~ covid_group * wave + age2 + sex + month_dif2 + month_dif2_sq + num_years_edu + eth2 + (1 | CC.ID), data = cattell_data2)
summary(model_cattell2)
cattell2_stand <- as.data.frame(standardize_parameters(model_cattell2))
# checking residuals
plot(resid(model_cattell1))
plot(qqnorm(resid(model_cattell1)))
plot(resid(model_cattell2))
plot(qqnorm(resid(model_cattell2)))
```
#boxplot
```{r}
ggplot(cattell_data, aes(x = wave, y = cattell_total_score, color = covid_group)) +
  geom_boxplot(size = 1) +
  labs(title = "Mean scores for instance with error bars", y = "Cattell score", x = "instance") +
  theme_cowplot(12) +
  theme(
    plot.title = element_blank(),
    axis.title = element_text(size = 20, face = "bold"), # originally 18 → now 20
    axis.text.x = element_text(size = 19), # originally 17 → now 19
    axis.text.y = element_text(size = 19),
    legend.title = element_text(size = 18, face = "bold"), # scale up from 16
    legend.text = element_text(size = 17), # scale up from 15
    strip.text = element_text(size = 18) # if using facet_wrap
  ) +
  scale_x_discrete(labels = c("T1", "T2", "T3")) +
  scale_color_manual(values = c("#999999", "#E69F00"), labels = c("Control group", "Covid group"), name = "Group")
```
#stw analysis
```{r}
# subselect relevant columns
stw_data <- df_A %>% select(CC.ID, wave, stw_n_correct_trials, covid_group, sex, ethnic_group, num_years_edu, stw_age, stw_date, eth2)

dates_wide <- stw_data %>%
  select(CC.ID, wave, stw_date) %>%
  distinct() %>% # in case of duplicates
  pivot_wider(
    names_from = wave, # e.g. "Phase1/2", "Phase4", ...
    values_from = stw_date,
    names_prefix = "stw_date_" # columns like story_memory_date_Phase1/2
  )
# calculate months difs between testing
stw_data <- stw_data %>%
  select(-stw_date) %>%
  left_join(dates_wide, by = "CC.ID")
stw_data <- stw_data %>%
  mutate(
    d_start = as.Date(`stw_date_Phase1/2`),
    d_middle = as.Date(stw_date_Phase4),
    d_end = as.Date(stw_date_Phase5),
    month_dif1 = time_length(interval(d_start, d_end), "months"),
    month_dif2 = time_length(interval(d_middle, d_end), "months")
  )

stw_data %>%
  filter(wave == "Phase5") %>%
  filter(covid_group == 0) %>%
  summary()

summary(stw_data$month_dif1, na.rm = TRUE)
summary(stw_data$month_dif2, na.rm = TRUE)

stw_data$month_dif1_sq <- stw_data$month_dif1^2
stw_data$month_dif2_sq <- stw_data$month_dif2^2
# calculate age at T2
stw_data$age2 <- ave(stw_data$stw_age, stw_data$CC.ID,
  FUN = function(x) if (all(is.na(x))) NA_real_ else max(x, na.rm = TRUE)
)
# check data types
str(stw_data)
stw_data$sex <- as.factor(cattell_data$sex)
stw_data$wave <- as.factor(cattell_data$wave)

# scale continuois columns
cols <- c("age2", "month_dif1", "month_dif1_sq", "month_dif2", "month_dif2_sq", "num_years_edu")
stw_data <- stw_data %>%
  dplyr::mutate(across(all_of(cols), ~ as.numeric(scale(.))))

contrasts(stw_data$wave)

dfwc <- summarySEwithin(stw_data,
  measurevar = "stw_n_correct_trials", withinvars = "wave", betweenvars = "covid_group",
  idvar = "CC.ID", na.rm = TRUE, conf.interval = .95
)


ggplot(dfwc, aes(x = wave, y = stw_n_correct_trials, group = covid_group, color = covid_group)) +
  geom_point(size = 4) +
  geom_line(size = 2) +
  geom_errorbar(aes(ymin = stw_n_correct_trials - se, ymax = stw_n_correct_trials + se), width = .1, size = 2) +
  theme_bw() +
  labs(title = "Mean scores for instance with error bars", y = "Stw score", x = "instance") +
  theme_cowplot(12) +
  theme(
    plot.title = element_blank(),
    axis.title = element_text(size = 20, face = "bold"), # originally 18 → now 20
    axis.text.x = element_text(size = 19), # originally 17 → now 19
    axis.text.y = element_text(size = 19),
    legend.title = element_text(size = 18, face = "bold"), # scale up from 16
    legend.text = element_text(size = 17), # scale up from 15
    strip.text = element_text(size = 18) # if using facet_wrap
  ) +
  scale_color_manual(values = c("#999999", "#E69F00"), labels = c("Control group", "Covid group")) +
  scale_x_discrete(labels = c("T1", "T2", "T3"))

stw_data$wave <- factor(stw_data$wave,
  levels = c("Phase1/2", "Phase4", "Phase5")
)
# dividing into 2 models

stw_data1 <- stw_data %>% filter(wave != "Phase4")

# model1
model_stw1 <- lmerTest::lmer(stw_n_correct_trials ~ covid_group * wave + age2 + sex + month_dif1 + month_dif1_sq + num_years_edu + eth2 + (1 | CC.ID), data = stw_data1)
summary(model_stw1)
standardize_parameters(model_stw1)
stw1_stand <- as.data.frame(standardize_parameters(model_stw1))
stw_data2 <- stw_data %>% filter(wave != "Phase1/2")
# model2
model_stw2 <- lmerTest::lmer(stw_n_correct_trials ~ covid_group * wave + age2 + sex + month_dif2 + month_dif2_sq + num_years_edu + eth2 + (1 | CC.ID), data = stw_data2)
summary(model_stw2)
standardize_parameters(model_stw2)
stw2_stand <- as.data.frame(standardize_parameters(model_stw2))


plot(resid(model_stw1))
plot(qqnorm(resid(model_stw1)))
plot(resid(model_stw2))
plot(qqnorm(resid(model_stw2)))
```
# correction for multiple comprisons - putting models together
```{r}
models <- list(model_mem1, model_mem2, model_acer, model_cattell1, model_cattell2, model_stw1, model_stw2)
model_names <- c("memory1", "memory2", "acer", "cattell1", "cattell2", "stw1", "stw2")  


# Extract fixed effects from all models
model_results <- lapply(seq_along(models), function(i) {
  mod <- models[[i]]
  res <- tidy(mod, effects = "fixed")
  res$model <- model_names[i]
  return(res)
})

# Combine all into one data frame
all_pvals <- do.call(rbind, model_results)
all_pvals
```

# correction for multiple comparisons - applying FDR
```{r}
fdr_pvals <- all_pvals %>%
  mutate(p_fdr = p.adjust(p.value, method = "fdr")) %>%
  ungroup()
options(scipen = 999)
fdr_pvals


fdr_pvals %>%
  filter(p_fdr <= 0.05) %>%
  select(term, model, p_fdr)
#summary file of all results
write.xlsx(fdr_pvals,"camcan_results.xlsx")
```
# descriptives
```{r}
df_A %>%
  group_by(wave, covid_group) %>%
  dplyr::summarise(
    dplyr::across(
      .cols = where(is.numeric) & !any_of(c("wave", "covid_group")),
      .fns = list(
        mean = ~ mean(.x, na.rm = TRUE),
        se   = ~ sd(.x, na.rm = TRUE) / sqrt(sum(!is.na(.x)))
      ),
      .names = "{.col}_{.fn}"
    ),
    .groups = "drop"
  )

dates <- demo %>%
  pivot_longer(
    cols = ends_with("_date"), # or: matches("_date$")
    names_to = "variable",
    values_to = "dates",
    values_drop_na = TRUE # optional: drop rows where dates are NA
  )


ggplot(dates, aes(x = dates, y = variable, colour = wave)) +
  geom_line() +
  theme_cowplot(12) +
  theme(
    plot.title = element_blank(),
    axis.title = element_text(size = 20, face = "bold"), # originally 18 → now 20
    axis.text.x = element_text(size = 19), # originally 17 → now 19
    axis.text.y = element_text(size = 19),
    legend.title = element_text(size = 18, face = "bold"), # scale up from 16
    legend.text = element_text(size = 17), # scale up from 15
    strip.text = element_text(size = 18) # if using facet_wrap
  ) +
  scale_color_manual(values = palette.colors(), labels = c("instance 1", "instance 2", "instance 3"))


cols <- c("stw_age", "acer_age", "story_memory_age", "cattell_age")

demo$age_sum <- rowSums(demo[cols], na.rm = TRUE)
demo$age_n <- rowSums(!is.na(demo[cols]))
demo$age_mean <- ifelse(demo$age_n > 0, demo$age_sum / demo$age_n, NA_real_)
demo$sex <- as.factor(demo$sex)
demo %>%
  filter(wave == "Phase5") %>%
  filter(covid_group == 1) %>%
  summary()
demo %>%
  filter(wave == "Phase1/2") %>%
  summary()
str(demo$sex)
df_A$eth2 <- as.factor(df_A$eth2)
df_A %>%
  filter(wave == "Phase5") %>%
  filter(covid_group == 0) %>%
  summary()
```
#results from other predictors table
```{r}

# Making sure list has names matching the object names
stand_mod <- list(
  mem1_stand      = mem1_stand,
  mem2_stand      = mem2_stand,
  acer_stand      = acer_stand,
  cattell1_stand  = cattell1_stand,
  cattell2_stand  = cattell2_stand,
  stw1_stand      = stw1_stand,
  stw2_stand      = stw2_stand
)

# Helper to format one df into: Parameter + "<dfname>" column
format_one <- function(df, nm) {
  df %>%
    transmute(
      Parameter,
      !!nm := sprintf("%.3f [%.3f, %.3f]", Std_Coefficient, CI_low, CI_high)
    )
}

# Apply to each df and join them wide by Parameter
formatted_list <- imap(stand_mod, format_one)

coef_wide <- reduce(formatted_list, full_join, by = "Parameter")

# View
coef_wide

#write_csv(coef_wide, "camcan_otherpred.csv")
```
